## Loading and preprocessing the data
```{r}
unzip("activity.zip")
data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

1. Make a histogram of the total number of steps taken each day

```{r}
# aggregate the data by steps per day
dates <- aggregate(steps ~ date, data = data, FUN = sum)
# plot it 
# using barplot as I can't figure out how to do properly x axis labelling
barplot(dates$steps, names.arg = dates$date, xlab = "date", ylab = "steps")
```

2. Calculate and report the **mean** and **median** total number of
steps taken per day

```{r}
mean(dates$steps)
median(dates$steps)
```
## What is the average daily activity pattern?

1. Make a time series plot (i.e. `type = "l"`) of the 5-minute
interval (x-axis) and the average number of steps taken, averaged
across all days (y-axis)

```{r}
# again, use the awesome aggregate method to collect data of steps based on interval
intervals <- aggregate(steps ~ interval, data = data, FUN = mean)
plot(intervals, type = "l")
```

2. Which 5-minute interval, on average across all the days in the
dataset, contains the maximum number of steps?

```{r}
# use the max method to index the intervals data frame to find the 5 minute
# interval with the maxiumum number of steps
intervals$interval[which.max(intervals$steps)]
```

## Imputing missing values

## Are there differences in activity patterns between weekdays and weekends?
