## Loading and preprocessing the data
```{r}
unzip("activity.zip")
data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

1. Make a histogram of the total number of steps taken each day

```{r}
# aggregate the data by steps per day
dates <- aggregate(steps ~ date, data = data, FUN = sum)
# plot it 
# using barplot as I can't figure out how to do properly x axis labelling
barplot(dates$steps, names.arg = dates$date, xlab = "date", ylab = "steps")
```

2. Calculate and report the **mean** and **median** total number of
steps taken per day

```{r}
mean(dates$steps)
median(dates$steps)
```
## What is the average daily activity pattern?

1. Make a time series plot (i.e. `type = "l"`) of the 5-minute
interval (x-axis) and the average number of steps taken, averaged
across all days (y-axis)

```{r}
# again, use the awesome aggregate method to collect data of steps based on interval
intervals <- aggregate(steps ~ interval, data = data, FUN = mean)
plot(intervals, type = "l")
```

2. Which 5-minute interval, on average across all the days in the
dataset, contains the maximum number of steps?

```{r}
# use the max method to index the intervals data frame to find the 5 minute
# interval with the maxiumum number of steps
intervals$interval[which.max(intervals$steps)]
```

## Imputing missing values

1. Calculate and report the total number of missing values in the
dataset (i.e. the total number of rows with `NA`s)

```{r}
# count how many values are missing
sum(is.na(activity))
```

2. Devise a strategy for filling in all of the missing values in the
dataset. The strategy does not need to be sophisticated. For
example, you could use the mean/median for that day, or the mean
for that 5-minute interval, etc.

I'll chose the second option and use the mean for the 5 minute interval.
I'll do it by merging in the intervals I just calculated and overwriting 
the NA values and then removing the merged interval column

3. Create a new dataset that is equal to the original dataset but with
the missing data filled in.

```{r}
# merge in the data and intervals
data <- merge(data, intervals, by = "interval", suffixes = c("",".data"))
# find the missing steps
missing <- is.na(data$steps)

# now write over teh NA's with teh data we got from the interval data frame
data$steps[missing] <- data$steps.data[missing]
# now rewrite data grabing columns 1 and 3, 
data <- data[,c(1:3)]
```

4. Make a histogram of the total number of steps taken each day and
Calculate and report the **mean** and **median** total number of
steps taken per day. Do these values differ from the estimates from
the first part of the assignment? 

```{r}
# very similar to step one, we use aggregate to get our data and then plot with barplot
intervals <- aggregate(steps ~ date, data = data, FUN = sum)
barplot(intervals$steps, names.arg = intervals$date, xlab="dates", ylab="steps per day")

mean(intervals$steps)
median(intervals$steps)
```

5. What is the impact of imputing missing data on the estimates of the total daily number of steps?

There does not seem to be any noticible impact.

## Are there differences in activity patterns between weekdays and weekends?
