## Loading and preprocessing the data
```{r}
unzip("activity.zip")
data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

1. Make a histogram of the total number of steps taken each day

```{r}
# aggregate the data by steps per day
dates <- aggregate(steps ~ date, data=data, FUN=sum)# plot it 
# using barplot as I can't figure out how to do properly x axis labelling
barplot(dates$steps, names.arg=dates$date, xlab="date", ylab="steps")
```

2. Calculate and report the **mean** and **median** total number of
steps taken per day

```{r}
mean(dates$steps)
median(dates$steps)
```
## What is the average daily activity pattern?

1. Make a time series plot (i.e. `type = "l"`) of the 5-minute
interval (x-axis) and the average number of steps taken, averaged
across all days (y-axis)

```{r}
# again, use the awesome aggregate method to collect data of steps based on interval
intervals <- aggregate(steps ~ interval, data=data, FUN=mean)
plot(intervals, type="l")
```

2. Which 5-minute interval, on average across all the days in the
dataset, contains the maximum number of steps?

```{r}
# use the max method to index the intervals data frame to find the 5 minute
# interval with the maxiumum number of steps
intervals$interval[which.max(intervals$steps)]
```

## Imputing missing values

1. Calculate and report the total number of missing values in the
dataset (i.e. the total number of rows with `NA`s)

```{r}
# count how many values are missing
sum(is.na(data))
```

2. Devise a strategy for filling in all of the missing values in the
dataset. The strategy does not need to be sophisticated. For
example, you could use the mean/median for that day, or the mean
for that 5-minute interval, etc.

I'll chose the second option and use the mean for the 5 minute interval.
I'll do it by merging in the intervals I just calculated and overwriting 
the NA values and then removing the merged interval column

3. Create a new dataset that is equal to the original dataset but with
the missing data filled in.

```{r}
# merge in the data and intervals
data <- merge(data, intervals, by="interval", suffixes=c("",".y"))
# find the missing steps
nas <- is.na(data$steps)

# now write over teh NA's with teh data we got from the interval data frame
data$steps[nas] <- data$steps.y[nas]

# now rewrite data grabing columns 1 and 3, 
data <- data[,c(1:3)]
```

4. Make a histogram of the total number of steps taken each day and
Calculate and report the **mean** and **median** total number of
steps taken per day. Do these values differ from the estimates from
the first part of the assignment? 

```{r}
# very similar to step one, we use aggregate to get our data and then plot with barplot
dates <- aggregate(steps ~ date, data=data, FUN=sum)
barplot(dates$steps, names.arg=dates$date, xlab="date", ylab="steps")

mean(dates$steps)
median(dates$steps)
```

5. What is the impact of imputing missing data on the estimates of the total daily number of steps?

There does not seem to be any noticible impact.

## Are there differences in activity patterns between weekdays and weekends?


```{r}
# we will use a function to determine if the current day is a weekend or weekday and label
# it accordingly with the new factors
daytype <- function(date) {
  if (weekdays(as.Date(date)) %in% c("Saturday", "Sunday")) {
    "weekend"
  } else {
    "weekday"
  }
}

data$daytype <- as.factor(sapply(data$date, daytype))

```

2. Make a panel plot containing a time series plot (i.e. `type = "l"`)
of the 5-minute interval (x-axis) and the average number of steps
taken, averaged across all weekday days or weekend days
(y-axis).

```{r}
# tell the device we'll have 1 row and two columns
par(mfrow=c(2,1))
for (type in c("weekend", "weekday")) {
  steps.type <- aggregate(steps ~ interval,
                          data=data,
                          subset=data$daytype==type,
                          FUN=mean)
  plot(steps.type, type="l", main=type)
}
```

